---
title: "Data visualization with ggplot2"
output: github_document
---

## How to get set up for this workshop

### Our virtual space:

[http://backchannelchat.com/chat/1ynme](http://backchannelchat.com/chat/1ynme)

### Getting this document on your computer:

1. Go to the GitHub repository here: [https://github.com/davidklaing/intro_to_ggplot2](https://github.com/davidklaing/intro_to_ggplot2)
2. Click the green button on the right that says "Clone or download".
3. Click "Download ZIP". (If you're proficient with git, feel free to fork the repository.)
4. Create a folder on your computer to store your work, and store your ZIP file there.
5. Double-click your ZIP file to unzip it and get all the code.

### Getting R and Rstudio

1. Download and install R from here: [http://cran.stat.sfu.ca/](http://cran.stat.sfu.ca/).
2. Download and install RStudio Desktop (Open Source Edition) from here: [https://www.rstudio.com/products/rstudio/#Desktop](https://www.rstudio.com/products/rstudio/#Desktop).

### Getting ready to play!

1. In RStudio, open `data_viz_with_ggplot2_workshop.Rmd`, a file in `YOUR_FOLDER/intro_to_ggplot2/`. (That's this file!)
2. In the code snippet below, remove the hashtags from both lines, and click the green "play" button on the right to install `dplyr`, `ggplot2`, and `titanic`, the three packages you'll need for the workshop. Once the installation is done, you can either delete those lines or comment them out (add hashtags).

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("gapminder")
```

### Rmarkdown in Rstudio: where the heck are we?

* RStudio is an Interactive Development Environment (or IDE) — a desktop application for writing code.
* Rmarkdown, the file format we're working in, allows us to write and execute code snippets, while also writing [markdown](https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet), a plain text format that allows for simple rendering of headings, bold, italics, hyperlinks, etc.
* Analogy: Rstudio is to Rmarkdown as Microsoft Word is to .docx.

Quick tip for writing and executing code in RStudio: to execute a line of code, move your cursor to that line and then type `Ctrl+Enter`. For example:

```{r}
# Move your cursor to the line below, and type Ctrl-Enter.
print("hello woRld!")
```

First we'll load our packages, using the `library()` function:

```{r}
library(dplyr)
library(ggplot2)
library(gapminder)
```

The `titanic` package has data about all the passengers of the Titanic. The two main datasets you can inspect are called `titanic_train` and `titanic_test` — the difference doesn't matter for now. Let's take a look at the first six rows of `gapminder`, using the `head()` function:

```{r}
head(gapminder)
```

What if we want to see some documentation on this dataset, perhaps to learn more about what the column names mean?

```{r}
# To see documentation about any object or function in R, prepend it with a 
# question mark and execute that line.
?gapminder
```

## Plotting with `ggplot()`

### Learning goals

1. Install and load the `ggplot2` package
2. Inspect a dataframe to see our data in spreadsheet form
3. Pass our dataframe into the `ggplot()` function, which builds the base layer for our visualization
4. Map our variables of interest onto the aesthetic properties of the graph, by using the `aes()` function in the `mapping` argument.
5. Define the geometric objects that we want to represent our data points (the rows in our dataframe)
6. Add a title, axis labels, and other goodies
7. Save a plot to PNG, for easy sharing




## Appendix

Plot for comparing data visualization options:

```{r}
viz_options <- tibble::tibble(
  option = c("ggplot2", "D3", "matplotlib", "Tableau"),
  simplicity = c(7, 1, 5, 10),
  flexibility = c(7, 10, 7, 4)
)
viz_options %>% 
  ggplot() +
  geom_point(aes(x = simplicity, y = flexibility)) +
  geom_text_repel(aes(x = simplicity, y = flexibility, label = option)) +
  scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  scale_y_continuous(limits = c(0, 10), breaks = 0:10) +
  labs(
    title = "My extremely subjective ratings of popular data viz tools",
    x = "How easy it is to learn",
    y = "How much you can do with it"
  ) +
  theme_light()
```

```{r}
sessionInfo()
```